---
title: "Analysis Predict Active Layer depth"
author: "Anna Talucci"
date: "2023-04-12"
output: html_document
---



# clear environment
```{r}
rm(list=ls())
```

# Overview

Standardize thaw depth measures to ALT. 


Needs to be cacluated by year because...

# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```

# Data
```{r}
data = read_csv("../data/era5Clean/ERA5CombineAll.csv")
```

```{r}
data
```
```{r}
( data1 = data %>% 
    drop_na(msrDOY)
)
```


## vector of data files
```{r}
f <- list.files(path = "../data/era5Clean/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```

## Individual files
```{r}
d2001 <- read_csv(f[1]) 
d2005 <- read_csv(f[2]) 
d2006 <- read_csv(f[3]) 
d2007 <- read_csv(f[4]) 
d2008 <- read_csv(f[5]) 
d2009 <- read_csv(f[6]) 
d2010 <- read_csv(f[7]) 
d2011 <- read_csv(f[8]) 
d2012 <- read_csv(f[9]) 
d2013 <- read_csv(f[10]) 
d2014 <- read_csv(f[11]) 
d2015 <- read_csv(f[12]) 
d2016 <- read_csv(f[13]) 
d2017 <- read_csv(f[14]) 
d2018 <- read_csv(f[15]) 
d2019 <- read_csv(f[16]) 
d2020 <- read_csv(f[17]) 
d2023 <- read_csv(f[18]) 
```

# Function for intial clean
```{r}
cleanFun = function(x) {
  x%>% mutate(altDate=minDate-14) %>% 
  rename(tempDOY = julianTemp, altDOY=altDate) %>%
  mutate(plotId=as.character((plotId)))-> x
  return(x)
}
```

```{r}
( f2001 = cleanFun(d2001) )
```

```{r}
f2005 = cleanFun(d2005)
f2006 = cleanFun(d2006)
f2007 = cleanFun(d2007)
f2008 = cleanFun(d2008)
f2009 = cleanFun(d2009)
f2010 = cleanFun(d2010)
```

```{r}
f2011 = cleanFun(d2011)
f2012 = cleanFun(d2012)
f2013 = cleanFun(d2013)
f2014 = cleanFun(d2014)
f2015 = cleanFun(d2015)
f2016 = cleanFun(d2016)
f2017 = cleanFun(d2017)
f2018 = cleanFun(d2018)
f2019 = cleanFun(d2019)
f2020 = cleanFun(d2020)
f2023 = cleanFun(d2023)
```


# Functions to standarize
Formula Parts 
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
B. Take the square root of the Sum of the total degrees above zero for dates prior to active layer thickness date ()

C. Divide B by A
F. Multiply thaw depth measure by C

```{r}
calc_A = function(x) {
  x %>% 
    group_by(id, msrDOY) %>% # group by id and measured date
    filter(tempDOY < msrDOY) %>% # filter DOY from ERA5 that are less then the measured Julian date
  filter(TempC > 0) %>% # filter for days over 0 degrees Celsius
    arrange(tempDOY) %>%
  summarize( 
  A_sum = sum(TempC),
  A = sqrt(A_sum)) -> x
  return(x)
}

```

**Change Date Depending on geographic location**
```{r}

calc_B = function(x) {
  x %>% 
    group_by(id, msrDOY) %>% # group by id and measured date
    filter(tempDOY < altDOY) %>% # filter Julian dates from ERA5 that are less then the alt date
  filter(TempC > 0) %>%  #filter for days over 0 degrees Celsius
    arrange(tempDOY) %>%
  summarize( 
  B_sum = sum(TempC),
  B = sqrt(B_sum)) -> x
  return(x)
 }
```

# Calculatations by each year

## Calculate Part A
```{r}
( calc_A_2001 = calc_A(f2001) )
```

```{r}
calc_A_2005 = calc_A(f2005) 
calc_A_2006 = calc_A(f2006) 
calc_A_2007 = calc_A(f2007) 
calc_A_2008 = calc_A(f2008) 
calc_A_2009 = calc_A(f2009) 
calc_A_2010 = calc_A(f2010) 
```

```{r}
calc_A_2011 = calc_A(f2011)
calc_A_2012 = calc_A(f2012) 
calc_A_2013 = calc_A(f2013) 
calc_A_2014 = calc_A(f2014) 
calc_A_2015 = calc_A(f2015) 
calc_A_2016 = calc_A(f2016) 
```

```{r}
calc_A_2017 = calc_A(f2017) 
calc_A_2018 = calc_A(f2018)  
calc_A_2019 = calc_A(f2019) 
calc_A_2020 = calc_A(f2020) 
calc_A_2023 = calc_A(f2023) 
```


## Calculate Part B
```{r}
( calc_B_2001 = calc_B(f2001) )
```

```{r}
calc_B_2005 = calc_B(f2005) 
calc_B_2006 = calc_B(f2006) 
calc_B_2007 = calc_B(f2007) 
calc_B_2008 = calc_B(f2008) 
calc_B_2009 = calc_B(f2009) 
calc_B_2010 = calc_B(f2010) 
```

```{r}
calc_B_2011 = calc_B(f2011)
calc_B_2012 = calc_B(f2012)
calc_B_2013 = calc_B(f2013)
calc_B_2014 = calc_B(f2014)
calc_B_2015 = calc_B(f2015)
calc_B_2016 = calc_B(f2016)
```

```{r}
calc_B_2017 = calc_B(f2017) 
calc_B_2018 = calc_B(f2018) 
calc_B_2019 = calc_B(f2019) 
calc_B_2020 = calc_B(f2020) 
calc_B_2023 = calc_B(f2023) 
```


## Filter original data 
```{r}
filterFun = function(x) {
  x %>% 
    dplyr::select(id, plotId:vegCvr) %>% # Select columns of interest for original data
  group_by(id) %>% # group by id
  slice(1)-> x #slice the data takes 1st row for each id
  return(x)
 }
```

```{r}
( orgf2001 = filterFun(f2001) )
```

```{r}
orgf2005 = filterFun(f2005) 
orgf2006 = filterFun(f2006) 
orgf2007 = filterFun(f2007) 
orgf2008 = filterFun(f2008) 
orgf2009 = filterFun(f2009) 
```

```{r}
orgf2010 = filterFun(f2010) 
orgf2011 = filterFun(f2011) 
orgf2012 = filterFun(f2012) 
orgf2013 = filterFun(f2013) 
orgf2014 = filterFun(f2014) 
```

```{r}
orgf2015 = filterFun(f2015) 
orgf2016 = filterFun(f2016) 
orgf2017 = filterFun(f2017) 
orgf2018 = filterFun(f2018) 
orgf2019 = filterFun(f2019) 
orgf2020 = filterFun(f2020) 
orgf2023 = filterFun(f2023) 
```

## Join Calculations with Orginial
```{r}
( data2001 = orgf2001 %>% 
    full_join(y=calc_A_2001, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2001, by=c("id", 'msrDOY')) )
```

```{r}
data2005 = orgf2005 %>% 
    full_join(y=calc_A_2005, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2005, by=c("id", 'msrDOY')) 
```

```{r}
data2006 = orgf2006 %>% 
    full_join(y=calc_A_2006, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2006, by=c("id", 'msrDOY')) 
```

```{r}
data2007 = orgf2007 %>% 
    full_join(y=calc_A_2007, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2007, by=c("id", 'msrDOY')) 
```


```{r}
data2008 = orgf2008 %>% 
    full_join(y=calc_A_2008, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2008, by=c("id", 'msrDOY')) 
```

```{r}
data2009 = orgf2009 %>% 
    full_join(y=calc_A_2009, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2009, by=c("id", 'msrDOY')) 
```

```{r}
data2010 = orgf2010 %>% 
    full_join(y=calc_A_2010, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2010, by=c("id", 'msrDOY')) 
```
```{r}
data2011 = orgf2011 %>% 
    full_join(y=calc_A_2011, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2011, by=c("id", 'msrDOY')) 
```

```{r}
data2012 = orgf2012 %>% 
    full_join(y=calc_A_2012, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2012, by=c("id", 'msrDOY')) 
```

```{r}
data2013 = orgf2013 %>% 
    full_join(y=calc_A_2013, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2013, by=c("id", 'msrDOY'))
```

```{r}
data2014 = orgf2014 %>% 
    full_join(y=calc_A_2014, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2014, by=c("id", 'msrDOY')) 
```

```{r}
data2015 = orgf2015 %>% 
    full_join(y=calc_A_2015, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2015, by=c("id", 'msrDOY')) 
```

```{r}
data2016 = orgf2016 %>% 
    full_join(y=calc_A_2016, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2016, by=c("id", 'msrDOY')) 
```

```{r}
data2017 = orgf2017 %>% 
    full_join(y=calc_A_2017, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2017, by=c("id", 'msrDOY')) 
```

```{r}
data2018 = orgf2018 %>% 
    full_join(y=calc_A_2018, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2018, by=c("id", 'msrDOY')) 
```

```{r}
data2019 = orgf2019 %>% 
    full_join(y=calc_A_2019, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2019, by=c("id", 'msrDOY')) 
```

```{r}
data2020 = orgf2020 %>% 
    full_join(y=calc_A_2020, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2020, by=c("id", 'msrDOY')) 
```

```{r}
data2023 = orgf2023 %>% 
    full_join(y=calc_A_2023, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2023, by=c("id", 'msrDOY')) 
```


## Combine standarized measures
```{r}
( 
data = dplyr::bind_rows(data2001, data2005, data2008, data2009, data2010, data2011, data2012, data2013, data2014, data2015, data2016, data2017, data2018, data2019, data2020)
)
```

## Finish prediction calculation
```{r}
( data1 = data %>%
    mutate(C = B/A) %>%
    mutate(predAlt = round((msrDepth*C), 0)) 
)
```

```{r}
( data_na = data1[!complete.cases(data1), ] )
```

```{r}
unique(data_na$year)
unique(data_na$lastNm)
```
## Save to csv
```{r}
write.csv(data1,file = "../outputs/Standardize/ThawStandardizeALT.csv")
```



## 2001
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2001
```
```{r}
f2001 %>% filter(id==1) %>% filter(tempDOY < msrDOY)
```
```{r}
sqrt(0)
```
```{r}
f2001 %>% 
    group_by(id) %>% # group by id and measured date
    filter(tempDOY < msrDOY) %>% # filter DOY from ERA5 that are less then the measured Julian date
  filter(TempC > 0) %>% # filter for days over 0 degrees Celsius
    arrange(tempDOY) %>%
  summarize( 
  A_sum = sum(TempC),
  A = sqrt(A_sum))
```

```{r}
f2001 %>% 
    group_by(id, msrDOY) %>% # group by id and measured date
    filter(tempDOY < altDOY) %>% # filter Julian dates from ERA5 that are less then the alt date
  filter(TempC > 0) %>%  #filter for days over 0 degrees Celsius
    arrange(tempDOY) %>%
  summarize( 
  B_sum = sum(TempC),
  B = sqrt(B_sum))
```
### Calculate A & B
```{r}
35.7/0

```




### Join Calculations with Orginial
```{r}
( data2001 = orgf2001 %>% 
    full_join(y=calc_A_2001, by=c("id", 'msrDOY')) %>%
    full_join(y=calc_B_2001, by=c("id", 'msrDOY')) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((msrDepth*C), 0)) 
)
```
### Check NAs
```{r}
data2001 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```


***********DELETE*******************
## 2005
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2005
f2005 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2005 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2005$id))
length(unique(f2005$last_nm))
```
### Calculate A & B


### Filter original data 


### Join Calculations with Orginial
```{r}
( data2005 = orgf2005 %>% 
    full_join(y=calc_A_2005, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2005, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0)) %>%
    mutate(plot_id = as.character(plot_id))
)
```
### Check NAs
```{r}
data2005 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2008
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2008
f2008 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2008 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2008$id))
length(unique(f2008$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2008 = calc_A(f2008) )
```

```{r}
( calc_B_2008 = calc_B(f2008) )
```

### Filter original data 
```{r}
( 
  orgf2008 = f2008 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2008 = orgf2008 %>% 
    full_join(y=calc_A_2008, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2008, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2008 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2009
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2009
f2009 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2009 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2009$id))
length(unique(f2009$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2009 = calc_A(f2009) )
```

```{r}
( calc_B_2009 = calc_B(f2009) )
```

### Filter original data 
```{r}
( 
  orgf2009 = f2009 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2009 = orgf2009 %>% 
    full_join(y=calc_A_2009, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2009, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2009 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2010
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2010
f2010 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2010 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2010$id))
length(unique(f2010$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2010 = calc_A(f2010) )
```

```{r}
( calc_B_2010 = calc_B(f2010) )
```

### Filter original data 
```{r}
( 
  orgf2010 = f2010 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2010 = orgf2010 %>% 
    full_join(y=calc_A_2010, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2010, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2010 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2011
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2011
f2011 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2011 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2011$id))
length(unique(f2011$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2011 = calc_A(f2011) )
```

```{r}
( calc_B_2011 = calc_B(f2011) )
```

### Filter original data 
```{r}
( 
  orgf2011 = f2011 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2011 = orgf2011 %>% 
    full_join(y=calc_A_2011, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2011, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2011 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```


## 2012
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2012
f2012 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2012 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2012$id))
length(unique(f2012$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2012 = calc_A(f2012) )
```

```{r}
( calc_B_2012 = calc_B(f2012) )
```

### Filter original data 
```{r}
( 
  orgf2012 = f2012 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2012 = orgf2012 %>% 
    full_join(y=calc_A_2012, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2012, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2012 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2013
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2013
f2013 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2013 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2013$id))
length(unique(f2013$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2013 = calc_A(f2013) )
```

```{r}
( calc_B_2013 = calc_B(f2013) )
```

### Filter original data 
```{r}
( 
  orgf2013 = f2013 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2013 = orgf2013 %>% 
    full_join(y=calc_A_2013, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2013, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2013 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2014
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2014
f2014 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2014 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2014$id))
length(unique(f2014$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2014 = calc_A(f2014) )
```

```{r}
( calc_B_2014 = calc_B(f2014) )
```

### Filter original data 
```{r}
( 
  orgf2014 = f2014 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2014 = orgf2014 %>% 
    full_join(y=calc_A_2014, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2014, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2014 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```



## 2015
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2015
f2015 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2015 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2015$id))
length(unique(f2015$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2015 = calc_A(f2015) )
```

```{r}
( calc_B_2015 = calc_B(f2015) )
```

### Filter original data 
```{r}
( 
  orgf2015 = f2015 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2015 = orgf2015 %>% 
    full_join(y=calc_A_2015, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2015, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2015 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2016
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2016
f2016 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2016 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2016$id))
length(unique(f2016$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2016 = calc_A(f2016) )
```

```{r}
( calc_B_2016 = calc_B(f2016) )
```

### Filter original data 
```{r}
( 
  orgf2016 = f2016 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2016 = orgf2016 %>% 
    full_join(y=calc_A_2016, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2016, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2016 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2017
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2017
f2017 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2017 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2017$id))
length(unique(f2017$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2017 = calc_A(f2017) )
```

```{r}
( calc_B_2017 = calc_B(f2017) )
```

### Filter original data 
```{r}
( 
  orgf2017 = f2017 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2017 = orgf2017 %>% 
    full_join(y=calc_A_2017, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2017, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2017 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2018
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2018
f2018 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2018 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2018$id))
length(unique(f2018$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2018 = calc_A(f2018) )
```

```{r}
( calc_B_2018 = calc_B(f2018) )
```

### Filter original data 
```{r}
( 
  orgf2018 = f2018 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2018 = orgf2018 %>% 
    full_join(y=calc_A_2018, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2018, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2018 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2019
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2019
f2019 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2019 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2019$id))
length(unique(f2019$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2019 = calc_A(f2019) )
```

```{r}
( calc_B_2019 = calc_B(f2019) )
```

### Filter original data 
```{r}
( 
  orgf2019 = f2019 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2019 = orgf2019 %>% 
    full_join(y=calc_A_2019, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2019, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2019 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

## 2020
### Checking some things
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
```{r}
f2020
f2020 %>% 
  group_by(id) %>%
  filter_at(vars(TempC, julianTemp, julinDt), all_vars(is.na(.)))

f2020 %>% 
    group_by(id) %>%
    filter(julianTemp < julinDt)

length(unique(f2020$id))
length(unique(f2020$last_nm))
```
### Calculate A & B
```{r}
( calc_A_2020 = calc_A(f2020) )
```

```{r}
( calc_B_2020 = calc_B(f2020) )
```

### Filter original data 
```{r}
( 
  orgf2020 = f2020 %>% 
  dplyr::select(id, plot_id:cntry_c, day:id) %>%
  group_by(id) %>%
  slice(1)
)
```

### Join Calculations with Orginial
```{r}
( data2020 = orgf2020 %>% 
    full_join(y=calc_A_2020, by=c("id", "julinDt")) %>%
    full_join(y=calc_B_2020, by=c("id", "julinDt")) %>%
    mutate(C = B/A) %>%
    mutate(AdjALT = round((thw_dpt*C), 0))
)
```
### Check NAs
```{r}
data2020 %>% filter_at(vars(A_sum, A, B_sum, B, C, AdjALT), all_vars(is.na(.)))
```

# Combine standarized measures
```{r}
( 
data = dplyr::bind_rows(data2001, data2005, data2008, data2009, data2010, data2011, data2012, data2013, data2014, data2015, data2016, data2017, data2018, data2019, data2020)
)
```

```{r}
write.csv(data,file = "../outputs/Standardize/ThawStandardizeALT.csv")
```