---
title: "ERA5 Data Organization"
author: "Anna Talucci"
date: "2023-03-25"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview
ERA5 data is extracted from Google Earth Engine and field sites by sample year. This script then takes that data and organizes it in a way that it can but sued to predict end of season active layer thickness.

Notes:
- 2016 was to large to process as one extraction in GEE; it is split by USA and Canada.

# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
library(data.table)
```

# Data

## Vector of data files
```{r}
f <- list.files(path = "../data/ERA5DataEE/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```

## Individual files
```{r}
f2001 <- read.csv(f[1], header = TRUE) 
f2005 <- read.csv(f[2], header = TRUE) 
f2006 <- read.csv(f[3], header = TRUE) 
f2007 <- read.csv(f[4], header = TRUE) 
f2008 <- read.csv(f[5], header = TRUE) 
f2009 <- read.csv(f[6], header = TRUE) 
f2010 <- read.csv(f[7], header = TRUE) 
f2011 <- read.csv(f[8], header = TRUE) 
f2012 <- read.csv(f[9], header = TRUE) 
f2013 <- read.csv(f[10], header = TRUE) 
f2014 <- read.csv(f[11], header = TRUE) 
f2015 <- read.csv(f[12], header = TRUE) 
f2016ca <- read.csv(f[13], header = TRUE) 
f2016us <- read.csv(f[14], header = TRUE) 
f2017 <- read.csv(f[15], header = TRUE) 
f2018 <- read.csv(f[16], header = TRUE) 
f2019 <- read.csv(f[17], header = TRUE) 
f2020 <- read.csv(f[18], header = TRUE) 
f2023 <- read.csv(f[19], header = TRUE) 
```

## View DF
```{r}
f2001
colnames(f2001)
```


## Process
1. Select columns of interest
1. pivot longer so column of temperature and date
2. Group by site_id
3. Select dates after day of thaw measurement
4. Select only rows with temp above 0 (zero)
5.  apply calculation

```{r}
( df2001 = f2001 %>%
  dplyr::select(biome:year, X20010401_temperature_2m:X20011031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2005 = f2005 %>%
  dplyr::select(biome:year, X20050401_temperature_2m:X20051031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2006 = f2006 %>%
  dplyr::select(biome:year, X20060401_temperature_2m:X20061031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2007 = f2007 %>%
  dplyr::select(biome:year, X20070401_temperature_2m:X20071031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2008 = f2008 %>%
  dplyr::select(biome:year, X20080401_temperature_2m:X20081031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2009 = f2009 %>%
  dplyr::select(biome:year, X20090401_temperature_2m:X20091031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2010 = f2010 %>%
  dplyr::select(biome:year, X20100401_temperature_2m:X20101031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2011 = f2011 %>%
  dplyr::select(biome:year, X20110401_temperature_2m:X20111031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2012 = f2012 %>%
  dplyr::select(biome:year, X20120401_temperature_2m:X20121031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2013 = f2013 %>%
  dplyr::select(biome:year, X20130401_temperature_2m:X20131031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2014 = f2014 %>%
  dplyr::select(biome:year, X20140401_temperature_2m:X20141031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2015 = f2015 %>%
  dplyr::select(biome:year, X20150401_temperature_2m:X20151031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
f2016 = bind_rows(f2016ca, f2016us)
( df2016 = f2016 %>%
  dplyr::select(biome:year, X20160401_temperature_2m:X20161031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2017 = f2017 %>%
  dplyr::select(biome:year, X20170401_temperature_2m:X20171031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2018 = f2018 %>%
  dplyr::select(biome:year, X20180401_temperature_2m:X20181031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2019 = f2019 %>%
  dplyr::select(biome:year, X20190401_temperature_2m:X20191031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
f2020
( df2020 = f2020 %>%
  dplyr::select(biome:year, X20200401_temperature_2m:X20201031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
f2023
( df2023 = f2023 %>%
  dplyr::select(biome:year, X20230401_temperature_2m:X20231031_temperature_2m) %>%
    mutate(id = row_number()) )
```

# Functions to clean 
```{r}
cleanDf = function(x)  {
  x %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(plotId = as.character(plotId)) %>%
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) %>%
    mutate(tempDate=ymd(tempDate)) %>%
    mutate(julianTemp = yday(tempDate)) -> x
  return(x)
}
```


# Apply functions to to each year
## Clean function
```{r}
( clean2001 = df2001 %>% cleanDf() )
```


```{r}
clean2005 = df2005 %>% cleanDf()
clean2006 = df2006 %>% cleanDf()
clean2007 = df2007 %>% cleanDf()
clean2008 = df2008 %>% cleanDf()
clean2009 = df2009 %>% cleanDf()
```
FYI the warning is referencing parts of the name being discarded in the separate process. This can be ignored for these chunks of code.
```{r}
clean2010 = df2010 %>% cleanDf()
clean2011 = df2011 %>% cleanDf()
clean2012 = df2012 %>% cleanDf()
clean2013 = df2013 %>% cleanDf()
clean2014 = df2014 %>% cleanDf()
```

```{r}
clean2015 = df2015 %>% cleanDf()
clean2016 = df2016 %>% cleanDf()
clean2017 = df2017 %>% cleanDf()
clean2018 = df2018 %>% cleanDf()
clean2019 = df2019 %>% cleanDf()
clean2020 = df2020 %>% cleanDf()
clean2023 = df2023 %>% cleanDf()
```
# Identify ALT date

## Functions to ID ALT Date
```{r}
altDateFun = function(x)  {
  x %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(julianTemp>=200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(lastNm, plotId, siteId, year) %>% # regroup by last name, plot id, and site id
  #filter(site_id== "DFTB") %>%
    arrange(julianTemp) %>% # sort by date of temp measures
  mutate(julianTemp_diff = julianTemp - lag(julianTemp)) %>%
  filter(julianTemp_diff >= 1) %>%
  summarize(minDate = min(julianTemp),
            meanDate = mean(julianTemp),
            maxDate = max(julianTemp)) %>%
    ungroup() %>%
  dplyr::select(lastNm, plotId, siteId, year, minDate) -> x
  return(x)
}
```


### 2001

```{r}
clean2001 
```

```{r}
(rm2001Fun = altDateFun(clean2001)) 
(rm2005Fun = altDateFun(clean2005)) 
(rm2006Fun = altDateFun(clean2006)) 
(rm2007Fun = altDateFun(clean2007)) 
(rm2008Fun = altDateFun(clean2008)) 
(rm2009Fun = altDateFun(clean2009)) 
```

```{r}
(rm2010Fun = altDateFun(clean2010)) 
(rm2011Fun = altDateFun(clean2011)) 
(rm2012Fun = altDateFun(clean2012)) 
(rm2013Fun = altDateFun(clean2013)) 
(rm2014Fun = altDateFun(clean2014)) 
(rm2015Fun = altDateFun(clean2015)) 
```

```{r}
(rm2016Fun = altDateFun(clean2016)) 
(rm2017Fun = altDateFun(clean2017)) 
(rm2018Fun = altDateFun(clean2018)) 
(rm2019Fun = altDateFun(clean2019)) 
(rm2020Fun = altDateFun(clean2020)) 
(rm2023Fun = altDateFun(clean2023)) 
```



# Combine the data
```{r}

combine2001 = merge(clean2001, rm2001Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2005 = merge(clean2005, rm2005Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

  combine2006 = merge(clean2006, rm2006Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

  combine2007 = merge(clean2007, rm2007Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 


  combine2008 = merge(clean2008, rm2008Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
 
  combine2009 = merge(clean2009, rm2009Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

  combine2010 = merge(clean2010, rm2010Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

```


```{r}
combine2011 = merge(clean2011, rm2011Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2012 = merge(clean2012, rm2012Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2013 = merge(clean2013, rm2013Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2014 = merge(clean2014, rm2014Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2015 = merge(clean2015, rm2015Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
```

```{r}
combine2016 = merge(clean2016, rm2016Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2017 = merge(clean2017, rm2017Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2018 = merge(clean2018, rm2018Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2019 = merge(clean2019, rm2019Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2020 = merge(clean2020, rm2020Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 

combine2023 = merge(clean2023, rm2023Fun[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
```

# write to csv 
```{r}
write.csv(combine2001,file = "../outputs/era5Clean/combine2001.csv")
write.csv(combine2005,file = "../outputs/era5Clean/combine2005.csv")
write.csv(combine2006,file = "../outputs/era5Clean/combine2006.csv")
write.csv(combine2007,file = "../outputs/era5Clean/combine2007.csv")
write.csv(combine2008,file = "../outputs/era5Clean/combine2008.csv")
write.csv(combine2009,file = "../outputs/era5Clean/combine2009.csv")
write.csv(combine2010,file = "../outputs/era5Clean/combine2010.csv")
write.csv(combine2011,file = "../outputs/era5Clean/combine2011.csv")
write.csv(combine2012,file = "../outputs/era5Clean/combine2012.csv")
write.csv(combine2013,file = "../outputs/era5Clean/combine2013.csv")
write.csv(combine2014,file = "../outputs/era5Clean/combine2014.csv")
write.csv(combine2015,file = "../outputs/era5Clean/combine2015.csv")
write.csv(combine2016,file = "../outputs/era5Clean/combine2016.csv")
write.csv(combine2017,file = "../outputs/era5Clean/combine2017.csv")
write.csv(combine2018,file = "../outputs/era5Clean/combine2018.csv")
write.csv(combine2019,file = "../outputs/era5Clean/combine2019.csv")
write.csv(combine2020,file = "../outputs/era5Clean/combine2020.csv")
write.csv(combine2023,file = "../outputs/era5Clean/combine2023.csv")
```

# Bind clean 
```{r}
( 
combineAll = dplyr::bind_rows(combine2001, combine2005, combine2006, combine2007, combine2008, combine2009, combine2010, combine2011, combine2012, combine2013, combine2014, combine2015, combine2016, combine2017, combine2018, combine2019, combine2020) %>%
  mutate(altDate = minDate-14)
)
```

## write to csv
```{r eval=FALSE, include=FALSE}
write.csv(combineAll,file = "../outputs/era5Clean/ERA5CombineAll.csv")
```



