---
title: "Figure Measured Vs Predicted"
author: "Anna Talucci"
date: "2023-05-18"
output: html_document
---


# clear environment
```{r}
rm(list=ls())
```

# Overview
Figures looking at measured versus predicted


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(ggnewscale)
```

# Plot themes
```{r}
comp_theme = theme_bw() + theme(legend.position = "none") +  
  theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"))
```

# Data

## Predicted measurements
```{r}
datacubePred = read_csv("../data/dataCube/DatacubePermafrostFirePredicted.csv")
```

```{r}
datacubePred
```

# Remove

```{r}
( datacubePred1 = datacubePred %>% filter(msrDepth < 1000) %>% drop_na(predDepth))
```

# Filter and sampling

## Small sample for paired figure
```{r}
( sampleActive = datacubePred %>% 
  filter(msrType =="active") %>%
  slice_sample(n = 20) %>%
    mutate(paired = row_number()) %>%
    dplyr::select(plotId, siteId, paired, msrType, msrDepth, predDepth) %>%
    pivot_longer(
    cols = ends_with("Depth"),
    names_to = "msrPrd",
    values_to = "depth")
    
)

```

## Data for Date figure
```{r}
( forDates = datacubePred %>% 
    dplyr::select(plotId, siteId, msrType, msrDOY, predDoy) %>%
    pivot_longer(
    cols = ends_with("doy"),
    names_to = "msrPrd",
    values_to = "DoY")
)
```

## Data for depth figure
```{r}
( forDepth = datacubePred %>% 
  dplyr::select(plotId, siteId, msrType, msrDepth, predDepth) %>%
    pivot_longer(
    cols = ends_with("Depth"),
    names_to = "msrPrd",
    values_to = "depth")
)
```


# Summarize
```{r}
forDates %>% 
  group_by(msrType, msrPrd) %>%
  summarise(mean= mean(date),
            sd = sd(date),
            min = min(date),
            max = max(date))
```

```{r}
forDepth %>% 
  group_by(thw_ctv, msrPrd) %>%
  summarise(mean= mean(depth),
            sd = sd(depth),
            min = min(depth),
            max = max(depth))
```

# Figures

## Violin plot of date measured vs predicted
```{r}
( fig1 = forDates %>% 
    filter(msrType=="thaw") %>% 
    mutate(msrPrd = fct_recode(msrPrd,
                                  "Measured DoY" = "msrDOY",
                                  "Predicted DoY" = "predDoy")) %>%
ggplot(., aes(x = msrPrd, y=DoY)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Day of year (DoY)") +
  comp_theme 
  
)
```

```{r}
( fig2 = forDates %>% 
    filter(msrType=="active") %>% 
    mutate(msrPrd = fct_recode(msrPrd,
                                  "Measured DoY" = "msrDOY",
                                  "Predicted DoY" = "predDoy")) %>%
ggplot(., aes(x = msrPrd, y=DoY)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  
  labs(x="", y="Day of year (DoY)") +
  comp_theme 
)
```

```{r}
( comboDates =  cowplot::plot_grid(fig1, fig2, ncol=1, labels = c('A', 'B'), label_size = 12)  ) 
```

```{r}
ggsave("../figures/DatesMeasurePredictViolin.png", plot = comboDates, width = 3, height =4, units = c("in"), dpi=600, bg = "white" )
```

## Violin plot of depth measured vs predicted
```{r}
( fig3 = forDepth %>% 
    filter(msrType=="thaw") %>% 
    mutate(msrPrd = fct_recode(msrPrd,
                                  "Measured depth" = "msrDepth",
                                  "Predicted depth" = "predDepth")) %>%
ggplot(., aes(x = msrPrd, y=depth)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  
  labs(x="", y="Depth (cm)") +
  comp_theme 
)
```

```{r}
( fig4 = forDepth %>% 
    filter(msrType=="active") %>% 
    mutate(msrPrd = fct_recode(msrPrd,
                                  "Measured depth" = "msrDepth",
                                  "Predicted depth" = "predDepth")) %>%
ggplot(., aes(x = msrPrd, y=depth)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +

  labs(x="", y="Depth (cm)") +
comp_theme 
)
```


```{r}
( comboDepths =  cowplot::plot_grid(fig3, fig4, ncol=1, labels = c('A', 'B'), label_size = 12)  ) 
```

```{r}
ggsave("../figures/DepthsMeasurePredictViolin.png", plot = comboDepths, width = 3, height =4, units = c("in"), dpi=600, bg = "white" )
```

## Observed vs predicted

```{r}
ggplot(data=datacubePred1, aes(x=msrDepth, y=predDepth)) +
  geom_point() +
  labs(x='Measured depth (cm)', y='Predicted depth (cm)', )
```

## Paired

```{r}
( paired2 = sampleActive %>% 
    mutate(msrPrd = fct_recode(msrPrd,
                                  "Measured active" = "dpthMsr",
                                  "Predicted ALT" = "dpthPrd")) %>%
    ggplot(., aes(x = msrPrd, y=depth)) +

  geom_line(aes(group = paired),  linetype = "dashed") +
  scale_color_manual(values = c('#7570b3', '#d95f02'), labels = c("Measured", "Predicted"), name="") +
     new_scale_color() +
        geom_point(aes(fill= msrPrd, color=msrPrd), size=3,  pch=21, stroke =1.8) +
          scale_fill_manual(name="", values=c( "#3b5a9d", "#FEC287FF"))  +
          scale_color_manual(name="", values = c( "grey79", "black")) +
  labs(x="", y="Depth (cm)") +
    comp_theme 


)

```

```{r}
ggsave("../figures/DepthsMeasurePredictActivePairedSample.png", plot = paired2, width = 3, height =3, units = c("in"), dpi=600, bg = "white" )
```