---
title: " Analysis Paired Burned Unburned"
author: "Anna Talucci"
date: "2024-02-23"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview
Figures looking at measured versus predicted


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(ggnewscale)
library(ggthemes)
```

# Plot themes
```{r}
comp_theme = theme_bw() + theme(legend.position = "bottom") +  
  theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"))
```

# Data

## Predicted measurements
```{r}
datacubePred = read_csv("../data/dataCube/PermafrostFireDatacubeFinal.csv")
```

```{r}
datacubePred
```
# Summary information

```{r}
unique(sort(datacubePred$fireYr))
```
```{r}
unique(sort(datacubePred$siteId))
```

```{r}
unique(sort(datacubePred$resName))
```

# Look at specific data
```{r}
datacubePred %>% filter(tsf==0)
```
```{r}
datacubePred1 = datacubePred %>%
  mutate(tsf = str_replace(tsf, 'unburned', 'u')) 
```
# aggregate burn and unburned pairs with summary stats

```{r}
( aggregate = datacubePred %>% 
    drop_na(paired) %>%
  group_by(resName, submitNm, distur, paired, fireYr, tsf) %>%
  summarise(n = n(),
            mean = mean(predDepth),
            sd   = sd(predDepth, na.rm=TRUE),
            se   = sd / sqrt(n))
)
```
```{r}
aggregate$tsf[aggregate$distur == "unburned"] <- "u"

```
```{r}
unique(sort(aggregate$tsf))
```

```{r}
aggregate$tsf = factor(aggregate$tsf, levels = c("u", "0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "12", "13", "14", "15", "16", "17", "18","19", "20", "21", "29", "30", "31", "32",  "33", "35", "36", "40", "41", "42", "45", "46", "47", "48", "49", "52"))
```



```{r}
unique(aggregate$tsf)
```



```{r}
aggregate 
```

```{r}
( aggregatePaired = datacubePred %>% 
    drop_na(paired) %>%
  group_by(resName, submitNm, distur, paired, fireYr) %>%
  summarise(n = n(),
            mean = mean(predDepth),
            sd   = sd(predDepth, na.rm=TRUE),
            se   = sd / sqrt(n))
)
```


# Plot

## Burned Unburned
```{r}
( pairedBoxplots = ggplot(aggregate, aes(x = distur, y=mean, color=distur)) +
  geom_boxplot() +
  scale_color_manual(values=c('#fc8d59', '#91bfdb'), name="", labels = c('Burned', 'Unburned')) +
  labs(x="", y="Depth (cm)") +
  facet_wrap(~resName, scales = "free_y") +
  comp_theme +
     theme(axis.text.y = element_text(size=5),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
    theme(
   strip.background = element_rect(
     color="white", fill="white", size=1.5, linetype="solid"
     )) +
    theme(strip.text.x = element_text(size = 5))
)
```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/AggregateBurnedUnburned.png", plot = pairedBoxplots, width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```

## Mean with Error bars by pairs By submit name
```{r}
pd <- position_dodge(0.5) # move them .05 to the left and right

( plotPairs = ggplot(aggregatePaired, aes(x = paired, y=mean, color=distur)) +
  #geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1, size=.3, position=pd) +
  geom_point(position=pd, size=1) +
  scale_color_manual(values=c('#fc8d59', '#91bfdb'), name="", labels = c('Burned', 'Unburned')) +
  labs(x="", y="Depth (cm)", ) +
 facet_wrap(~submitNm, scales = "free", ncol=4) +
  comp_theme +
    theme(axis.text.y = element_text(size=5),
          axis.text.x = element_text(size=5, angle=45),
   strip.background = element_rect(color="white", fill="white", size=1.5, linetype="solid"),
  strip.text.x = element_text(size = 8)) 
)

```
width = 6, height =5, units = c("in"),
```{r eval=FALSE, include=FALSE}
ggsave("../figures/AggregateBurnedUnburnedPairs.png", plot = plotPairs,  dpi=600, bg = "white" )
```

# Means with error Bars with tsf

```{r}
( plotMeanTSF <- ggplot(aggregate, aes(x = tsf, y=mean, color=distur, group=tsf)) +
    geom_point() +
      scale_color_manual(values=c('#fc8d59', '#91bfdb'), name="", labels = c('Burned', 'Unburned')) +
  labs(x="Time since fire", y="Depth (cm)") +
  facet_wrap(~resName, scales = "free") +
  comp_theme +
    theme(
          axis.text.y = element_text(size=5),
          axis.text.x = element_text(size=5, angle=45),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 6))
)
```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/AggregateBurnedUnburnedTsf.png", plot = plotMeanTSF, width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```

# MEan TSF repeat measures

```{r}
aggregate1 = aggregate %>% filter(paired %in% c("a1", "a2", "a4", "f1", "g1", "i2", "i1", "l2", "l3", "l4", "l6", "n7", "n1", "n2", "n4", "n5", "p1", "p6"))

```

```{r}
aggregate1
```

```{r}
( plotMeanTSFrepeat <- ggplot(aggregate1, aes(x = tsf, y=mean, color=submitNm)) +
    geom_point() +
      scale_color_manual(values=c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02','#a6761d'), name="") +
  labs(x="Time since fire", y="Depth (cm)") +
  facet_wrap(~resName, scales = "free") +
  comp_theme +
    theme(
          axis.text.y = element_text(size=5),
          axis.text.x = element_text(size=5, angle=45),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 6))
)
```


## Mean with Error bars by pairs By submit name

```{r}
unique(aggregate$submitNm)
```

```{r}
pd <- position_dodge(0.5) # move them .05 to the left and right
```

```{r}
aggregate %>% filter(submitNm == "Baillargeon") %>%
  ggplot(., aes(x = paired, y=mean, color=distur)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.3, position=pd) +
  geom_point(position=pd, size=2) +
  scale_color_manual(values=c('#fc8d59', '#91bfdb'), name="", labels = c('Burned', 'Unburned')) +
  labs(x="Pairs", y="Depth (cm)", title="Baillargeon") +
  comp_theme +
    theme(
   strip.background = element_rect(
     color="white", fill="white", size=1.5, linetype="solid"
     )) +
    theme(strip.text.x = element_text(size = 6)) 
 

```

```{r}
aggregate %>% filter(submitNm == "Gaglioti" ) %>%
  ggplot(., aes(x = paired, y=mean, color=distur)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.3, position=pd) +
  geom_point(position=pd, size=2) +
  scale_color_manual(values=c('#fc8d59', '#91bfdb'), name="", labels = c('Burned', 'Unburned')) +
  labs(x="Pairs", y="Depth (cm)", title="Gaglioti") +
  comp_theme +
    theme(
   strip.background = element_rect(
     color="white", fill="white", size=1.5, linetype="solid"
     )) +
    theme(strip.text.x = element_text(size = 6)) 
```

```{r}
aggregate %>% filter(submitNm == "Natali" ) %>%
  ggplot(., aes(x = paired, y=mean, color=distur)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.3, position=pd) +
  geom_point(position=pd, size=2) +
  scale_color_manual(values=c('#fc8d59', '#91bfdb'), name="", labels = c('Burned', 'Unburned')) +
  labs(x="Pairs", y="Depth (cm)", title="Natali") +
  comp_theme +
    theme(
   strip.background = element_rect(
     color="white", fill="white", size=1.5, linetype="solid"
     )) +
    theme(strip.text.x = element_text(size = 6)) 
```

```{r}
aggregate %>% filter(submitNm == "Rocha" ) %>%
  ggplot(., aes(x = paired, y=mean, color=distur)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.3, position=pd) +
  geom_point(position=pd, size=2) +
  scale_color_manual(values=c('#fc8d59', '#91bfdb'), name="", labels = c('Burned', 'Unburned')) +
  labs(x="Pairs", y="Depth (cm)", title="Rocha") +
  comp_theme +
    theme(
   strip.background = element_rect(
     color="white", fill="white", size=1.5, linetype="solid"
     )) +
    theme(strip.text.x = element_text(size = 6)) 
```

```{r}
aggregate %>% filter(submitNm == "Rocha" ) %>%
  ggplot(., aes(x = paired, y=mean, color=distur)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.3, position=pd) +
  geom_point(position=pd, size=2) +
  scale_color_manual(values=c('#fc8d59', '#91bfdb'), name="", labels = c('Burned', 'Unburned')) +
  labs(x="Pairs", y="Depth (cm)", title="Rocha") +
  comp_theme +
    theme(
   strip.background = element_rect(
     color="white", fill="white", size=1.5, linetype="solid"
     )) +
    theme(strip.text.x = element_text(size = 6)) 
```