---
title: "PredictALT"
author: "Anna Talucci"
date: "2024-02-13"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview
Take field measure active layer depth and predict maximum depth based on air temperature.

# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```

# Data

```{r}
data = read_csv("../data/dataCube/DatacubePermafrostFireERA5.csv")
```

```{r}
glimpse(data)
```
```{r}
( manies = data %>% filter(lastNm=="Manies") %>%filter(id <=41) )
unique(manies$msrDOY)
```

```{r}
( natali = data %>% filter(lastNm=="Natali") %>% filter(year ==2015) %>% filter(id>=3365) %>% filter(id<=3484)) %>% filter(TempDoy>259)

```
3365- 3484
# Functions to standarize
Formula Parts 
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
B. Take the square root of the Sum of the total degrees above zero for dates prior to active layer thickness date ()

C. Divide B by A
F. Multiply thaw depth measure by C

```{r}
calc_A = function(x) {
  x %>% 
    group_by(year, id, lastNm, msrDOY) %>%
    filter(TempDoy < msrDOY) %>%
  filter(TempC > 0) %>%
    arrange(TempDoy) %>%
  summarize( 
  A_sum = sum(TempC),
  A = sqrt(A_sum)) -> x
  return(x)
}

```

**Change Date Depending on geographic location**
```{r}

calc_B = function(x) {
  x %>% 
    group_by(year, id, lastNm, msrDOY) %>%
    filter(TempDoy < predDoy) %>%
  filter(TempC > 0) %>%
    arrange(TempDoy) %>%
  summarize( 
  B_sum = sum(TempC),
  B = sqrt(B_sum)) -> x
  return(x)
 }
```



### Calculate A & B
```{r}
( calc_A_data = calc_A(data) )
```

```{r}
( calc_B_data = calc_B(data) )
```

## Figure out na data in Manies id 1-41
```{r}
calc_A_data %>% filter(lastNm=="Manies") %>% filter(id <=41)
```

```{r}
manies %>% 
    group_by(year, id, lastNm, msrDOY) %>%
    filter(TempDoy < msrDOY) %>%
  filter(TempC > 0) %>%
    arrange(TempDoy) %>% 
  filter(id <=41)

```
3365- 3484
### Filter original data 
```{r}
( 
  orgData = data %>% 
  dplyr::select(plotId:predDoy) %>%
  group_by(year, id, lastNm) %>%
  slice(1) %>%
    dplyr::select(plotId:ECO_NAME, minDate, predDoy)
)
```

### Join Calculations with Orginial
```{r}
( dataCombine = orgData %>% 
    full_join(y=calc_A_data, by=c("year", "id", 'lastNm', 'msrDOY')) %>%
    full_join(y=calc_B_data, by=c("year", "id", 'lastNm', 'msrDOY')) %>%
    mutate(A_sum = ifelse(is.na(A_sum), 0, A_sum),
           A =ifelse(is.na(A), 0, A)) %>%
    mutate(C = B/A) %>%
    mutate(predDepth = round((msrDepth*C), 0)) 
    )
```


### Write to csv

```{r}
write.csv(dataCombine, file = "../outputs/dataCube/DatacubePermafrostFirePredicted.csv", row.names=FALSE)
```

