---
title: "QAQC Datacube"
author: "Anna Talucci"
date: "2024-01-30"
output: html_document
---


# clear environment
```{r}
rm(list=ls())
```

# Overview
QAQC of full contributed data set. Identify NAs for key variables (e.g., DOY, msrDepth, hitRock, gtProbe).


# Packages

```{r}
library(tidyverse)
library(lubridate)
```

# Data

```{r}
datacube = read_csv("../data/dataCube/PermafrostFireDataCube.csv")
```

```{r}
datacube
```

```{r}
datacubePred = read_csv("../data/dataCube/DatacubePermafrostFirePredicted.csv")
```

# Existing NAs

## Check missing DOY
As of 1-31-2024 Dieleman is missing day of month for 1843 measurements
```{r}
( 
  DOY_na = datacube %>% filter_at(vars(msrDOY), all_vars(is.na(.))) 
)

unique(DOY_na$lastNm)
```

## Drop NA
```{r eval=FALSE, include=FALSE}
(
  dropDOYNA = datacube %>% filter(msrDOY>0)
)
```

## missing measurements
```{r}
(msrNA = datacube %>% filter(is.na(msrDepth)) )
unique(msrNA$lastNm)
```
```{r}
(rocky = datacube %>% filter(is.na(msrDepth)) %>% filter(hitRock=="y") )
unique(rocky$lastNm)
```

```{r}
(rockno = datacube %>% filter(is.na(msrDepth)) %>% filter(hitRock=="n") )
unique(rockno$lastNm)
```
```{r}
( naMsrDepth = datacube %>% filter(is.na(msrDepth)) )
```
```{r}
unique(naMsrDepth$submitNm)
```

```{r}
naMsrDepth %>% filter(submitNm == "Baillargeon") # missing msrDepth n=8 inundation
naMsrDepth %>% filter(submitNm == "Loranty") # missing msrDepth n=9 Hit rock
naMsrDepth %>% filter(submitNm == "Natali") # missing msrDepth n=24 missing in submitted data
naMsrDepth %>% filter(submitNm == "O'Donnell") # missing msrDepth n=5
naMsrDepth %>% filter(submitNm == "Rocha") # missing msrDepth n=390 missing in submitted data
naMsrDepth %>% filter(submitNm ==  "Sizov") # missing msrDepth n=4 missing in submitted data
```


```{r}
datacube %>% filter(is.na(msrDepth)) %>% filter(hitRock=="n") %>% filter(submitNm=="Baillargeon")
datacube %>% filter(is.na(msrDepth)) %>% filter(hitRock=="n") %>% filter(submitNm=="Natali")

datacube %>% filter(is.na(msrDepth)) %>% filter(hitRock=="n") %>% filter(submitNm=="O'Donnell")
datacube %>% filter(is.na(msrDepth)) %>% filter(hitRock=="n") %>% filter(submitNm=="Rocha")

datacube %>% filter(is.na(msrDepth)) %>% filter(hitRock=="n") %>% filter(submitNm=="Sizov")
```
For Baillargeon, the NA values for measurements (n=8) were a function of an existing transect subsiding into a lake.
For Natali, the NA values for measurements (n=41) were a function ....
For O'Donnell, the NA values for measurements (n=4) were a function ....
For Rocha, the NA values for measurements (n=390) were a function ....
For Sizov, the NA values for measurements (n=4) were a function ....


## HitRock yes 

```{r}
(yesRock = datacube %>% filter(hitRock=="y") )
unique(yesRock$lastNm)
```
Measurements that hit Rock n=882

```{r}
(gtProbe = datacube %>% filter(gtProbe=="y") )
unique(gtProbe$lastNm)
```

Measurements greater than the measurement probe n=2225

# Summary characteristics for tables
```{r}
datacube %>% group_by(lastNm, cntryId, biome) %>%
  summarize(fireYears = paste(sort(unique(fireYr)), collapse = ", ")) %>%
  write_csv(., '../outputs/summaryTables/SummaryFireYear.csv')
```

# Counts
```{r}
datacube %>% group_by(cntryId, biome, siteId, lastNm) %>%
  tally()
```

```{r}
datacube %>% group_by(cntryId, biome,  plotId, siteId, lastNm) %>%
  summarise(n=n())
```

```{r}
datacube %>% group_by(cntryId, biome,  fireYr, lastNm ) %>%
  summarize(n=n())
```

```{r}
datacube %>% group_by(lastNm ) %>%
  summarize(n=n())
```

```{r}
datacube %>% filter(lastNm=="Schulze")
```

# Predicted Data

```{r}
datacubePred
```

```{r}
( naPredDepth = datacubePred %>% filter(is.na(predDepth)) )
```

```{r}
unique(naPredDepth$lastNm)
```


```{r}
naPredDepth %>% filter(lastNm=="Manies") # sum_A not calculating correctly need to check for id 1-41 because the temperatures for the days preceeding 117 are all below freezing
naPredDepth %>% filter(lastNm=="O'Donnell") # missing msrDepth n=5 missing in submitted
naPredDepth %>% filter(lastNm=="Rocha") # missing msrDepth n= 390 missing in submitted
naPredDepth %>% filter(lastNm=="Loranty") # missing msrDepth n=9 missing in submitted
naPredDepth %>% filter(lastNm=="Dieleman_Baltzer_Turetsky") # missing day of month n=1841
naPredDepth %>% filter(lastNm=="Natali") # Something weird with the Temperature date from ERA5 need to check n=237 on missing n=24 msrDepth in submitted data
naPredDepth %>% filter(lastNm=="Baillargeon, MacArthur, Natali, Sistla") # missing msrDepth n=8 missing in submitted
naPredDepth %>% filter(lastNm=="Sizov, Soromotin, Brodt") # missing msrDepth n=4 missing in submitted
```
3365- 3484	
## Large depth values

```{r}
datacubePred %>% filter(msrDepth > 300)

```

```{r}
datacubePred %>% filter(msrDepth < 0)

```

# Filter out data with errors
This includes a negative thaw depth

Note from Olefeldt --  all three look like typos. Our thaw rod only is 150 cm, so all numbers need to be within that. 
```{r}
datacubePred
```
```{r}
( 
  datacubePredClean = datacubePred %>% 
  filter(!msrDepth < 0) %>% # Remove Olefeldt negative measurements (also removes msrDepth with NA)
  filter(!msrDepth %in% c(4501, 453, 1032))
)
```

%>% 
    filter(msrDepth < 1000) %>% 
    drop_na(predDepth) %>% 
    filter_at(vars(predDepth), all_vars(!is.infinite(.)))

```{r}
setdiff( datacubePred, datacubePredClean)
```

```{r}
datacubePredClean %>% filter(is.na(predDepth))
```

```{r eval=FALSE, include=FALSE}
write.csv(datacubePredClean, file = "../outputs/dataCube/datacubePredClean.csv", row.names = FALSE)
```

**THE END**