---
title: "Datacube spatial adds"
author: "Anna Talucci"
date: "2024-01-31"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```


# Overview
Create ecoregion centroids for Manuscript map.

Add Ecozones to points.



Notes:
[Better colors for Mapping](https://www.esri.com/arcgis-blog/products/js-api-arcgis/mapping/better-colors-for-better-mapping/)

Publically available data for download used here:
[Resolve Ecoregions](https://ecoregions.appspot.com/)
[Resolve Ecoregions for download](https://hub.arcgis.com/datasets/37ea320eebb647c6838c23f72abae5ef/explore)

[WWF Ecoregions](https://www.arcgis.com/apps/View/index.html?appid=d60ec415febb4874ac5e0960a6a2e448)
[WWF EcoRegions download](https://www.worldwildlife.org/publications/terrestrial-ecoregions-of-the-world)
# Packages
```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
library(ggnewscale)
```

# Projection
[Some projection info]https://nsidc.org/data/user-resources/help-center/guide-nsidcs-polar-stereographic-projection)

```{r}
polarProj = "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=-45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
```

# Data

## Points

```{r}
datacubePts = st_read("../data/dataCube/PermafrostFireDataCube.shp", "PermafrostFireDataCube")
```

## WWF
```{r}
wwf = st_read("../data/spatialData/wwfEcozones/wwf_terr_ecos.shp", "wwf_terr_ecos")
```

```{r}
wwf 
```
```{r}
wwf %>% st_drop_geometry() %>% distinct(REALM)
```


## Resolve
```{r}
resolve = st_read("../data/spatialData/RESOLVE_Ecoregions_and_Biomes/Biomes_and_Ecoregions_2017.shp", "Biomes_and_Ecoregions_2017")
```

```{r}
resolve
```
## Permafrost

```{r}
permafrost = st_read("../data/spatialData/permafrost/permaice.shp", "permaice")
```

```{r}
permafrost
```

```{r}
permafrost_wgs = st_transform(permafrost, crs= st_crs(datacubePts))
```

```{r}
permafrost_wgs
```

```{r}
plot(permafrost_wgs)
```

```{r}
unique(permafrost_wgs$NUM_CODE)
unique(permafrost_wgs$COMBO)
unique(permafrost_wgs$RELICT)
unique(permafrost_wgs$EXTENT)
unique(permafrost_wgs$CONTENT)
unique(permafrost_wgs$LANDFORM)
```

# Filter for WWF Ecoregions
```{r}
( 
  tunBorWwf = wwf %>% 
    filter(REALM %in% c("PA", "NA")) %>% 
    dplyr::select(OBJECTID, ECO_NAME, REALM, BIOME, ECO_NUM, ECO_ID, geometry) %>% 
    filter(BIOME %in% c(11, 6)) %>%
    rename(wwfId = OBJECTID, wwfName=ECO_NAME, wwfRealm = REALM, wwfBiome = BIOME, wwfNum = ECO_NUM, wwfEcoId = ECO_ID)
)
```



```{r}
wwfPt = st_join(datacubePts, tunBorWwf) 
```

```{r}
names(wwfPt)
```
```{r}
(diaz = wwfPt %>% filter(submitNm=="Diaz") )

unique(diaz$wwfName)
```

## Add Resolve ecozones
```{r}
( resTunBor = resolve %>% 
    filter(REALM %in% c("Palearctic", "Nearctic")) %>% 
    dplyr::select(OBJECTID, ECO_NAME, BIOME_NUM, BIOME_NAME, REALM, ECO_ID, geometry) %>% 
    filter(BIOME_NUM %in% c(11, 6)) %>%
    rename(resId = OBJECTID, resName=ECO_NAME, resBiomNm = BIOME_NUM, resBiomeName = BIOME_NAME, resRealm = REALM,  resEcoId = ECO_ID)
)
```

```{r}
resWwfPt = st_join(wwfPt, resTunBor) 
```
```{r}
resWwfPt
```
```{r}
(diaz = resWwfPt %>% filter(submitNm=="Diaz") )

unique(diaz$resName)
unique(diaz$wwfName)
```

```{r}
resWwfPt %>% filter(is.na(resName)) 
```

```{r}
resWwfPt %>% filter(is.na(wwfName)) 
```

```{r}
resWwfPt %>% filter(is.na(msrDoy)) 
```

## Add Permafrost extent
```{r}
resWwfPt_prj = st_transform(resWwfPt, crs=st_crs(permafrost))
```

```{r}
resWwfPt_prj
```





```{r}
permResWwfPt = st_join(resWwfPt_prj, permafrost) 
```

```{r}
( permResWwfPtRename = permResWwfPt %>% 
  rename(permaCode = NUM_CODE, permaCombo =COMBO , permaRelict =RELICT , permaExtent= EXTENT, permaCont = CONTENT, permaLand = LANDFORM) )
```

```{r}
permResWwfPt_Wgs = st_transform(permResWwfPtRename, crs=4326)
```

```{r}
permResWwfPt_Wgs
```

## write to shapefile
```{r eval=FALSE, include=FALSE}
st_write(permResWwfPt_Wgs, "../outputs/spatialDataCreate/PermafrostFireEcozoneDatacube.shp", driver="ESRI Shapefile")
```



```{r eval=FALSE, include=FALSE}
st_write(diaz, "../outputs/spatialDataCreate/diaz.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(tunBorWwf, "../outputs/spatialDataCreate/WwfEcoregionsTunBor.shp", driver="ESRI Shapefile")
```

```{r}
unique(resWwfPt$wwfName)
```

```{r}
cat(paste0(sprintf('"%s"', sort(unique(resWwfPt$wwfName))), collapse = ", "))
```

# WWF Ecozones for Map

## Summaries sites by 

```{r}
(
  wwfPtCt <- resWwfPt %>% st_drop_geometry() %>%
                group_by(wwfName) %>%
                summarise(count = n()) 
  )
```
```{r}
( wwfPtSum <- resWwfPt %>%
                group_by(wwfName) %>%
                summarise(st_union(geometry)) %>%
                st_centroid() %>%
  left_join(., wwfPtCt, by= 'wwfName') %>%
  mutate(total = sum(count)) %>%
  mutate(prop = round(count/total, 5)) %>%
  mutate(percent = round(prop*100, 2)) %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry() %>%
    drop_na(wwfName) %>%
  mutate(prct = ifelse(percent == 0.38, 0.4,
                ifelse(percent == 43.24, 43.2,
                ifelse(percent == 7.01, 7.0,
                ifelse(percent == 0.08, 0.1,
                ifelse(percent == 0.96, 1.0,
                ifelse(percent == 0.34, 0.3,
                ifelse(percent == 17.45, 17.5,
                ifelse(percent == 1.68, 1.7,
                ifelse(percent == 22.3, 22.3,
                ifelse(percent == 1.33, 1.3,
                ifelse(percent == 2.15, 2.2,
                ifelse(percent == 3.06, 3.1,
                ifelse(percent == 0.01, 0.01, -99))))))))))))))
  )

```

```{r}
write_csv(wwfPtSum, '../outputs/wwfCentroidSummary.csv')
```

## Centroid by each region
Fore the overview map
### 1. "Arctic coastal tundra"
```{r}
( wwf01 =  tunBorWwf %>% 
    filter(wwfName=='Arctic coastal tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Arctic coastal tundra')
)
```

### 2. "Arctic foothills tundra"
```{r}
wwf02 =  tunBorWwf %>% 
    filter(wwfName=='Arctic foothills tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Arctic foothills tundra')
```

### 3. "Beringia lowland tundra"
```{r}
wwf03 =  tunBorWwf %>% 
    filter(wwfName=='Beringia lowland tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Beringia lowland tundra')
```


3              Beringia lowland tundra POINT (-160.8539 60.86536)
4               Beringia upland tundra POINT (-161.7837 62.63477)

### 4. "Beringia upland tundra" 
```{r}
( wwf04 =  tunBorWwf %>% 
    filter(wwfName=='Beringia upland tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Beringia upland tundra') )
```
Longitude: -165.43798
Latitude: 64.52649
### 5. Chukchi Peninsula tundra 
```{r}
wwf05 =  tunBorWwf %>% 
    filter(wwfName=="Chukchi Peninsula tundra") %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName="Chukchi Peninsula tundra")
```

### 6. "East Siberian taiga" 
```{r}
wwf06 =  tunBorWwf %>% 
    filter(wwfName=="East Siberian taiga") %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName="East Siberian taiga")
```


### 7. "Interior Alaska-Yukon lowland taiga"
```{r}
wwf07 =  tunBorWwf %>% 
    filter(wwfName=='Interior Alaska-Yukon lowland taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Interior Alaska-Yukon lowland taiga')
```

### 8. "Interior Yukon-Alaska alpine tundra"
```{r}
wwf08 =  tunBorWwf %>% 
    filter(wwfName=='Interior Yukon-Alaska alpine tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Interior Yukon-Alaska alpine tundra')
```

### 9. "Muskwa-Slave Lake forests"
```{r}
wwf09 =  tunBorWwf %>% 
    filter(wwfName=='Muskwa-Slave Lake forests') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Muskwa-Slave Lake forests')
```

### 10. "Northeast Siberian taiga"
```{r}
wwf10=  tunBorWwf %>% 
    filter(wwfName=='Northeast Siberian taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Northeast Siberian taiga')
```

### 11. "Northern Canadian Shield taiga"
```{r}
wwf11=  tunBorWwf %>% 
    filter(wwfName=='Northern Canadian Shield taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Northern Canadian Shield taiga')
```
### 12. "Northwest Territories taiga"
```{r}
( wwf12=  tunBorWwf %>% 
    filter(wwfName=='Northwest Territories taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName='Northwest Territories taiga')
)
```

### 13. "Yamal-Gydan tundra"
```{r}
wwf13 =  tunBorWwf %>% 
    filter(wwfName=="Yamal-Gydan tundra") %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(wwfName="Yamal-Gydan tundra")
```

## Recombine WWF ecoregions
```{r}
( 
  wwfCentroid = bind_rows(wwf01, wwf02, wwf03, wwf04, wwf05, wwf06, wwf07, wwf08, wwf09, wwf10, wwf11, wwf12, wwf13) %>% 
    rename(geometry = x)
)
```
 
## Edit geometry of Beringa Upland and lowland to limit overlap
```{r}
wwfCentroid
```
Longitude: -165.43798
Latitude: 64.52649
```{r}
st_geometry(wwfCentroid) <- st_sfc(
    ifelse(wwfCentroid$wwfName == "Beringia upland tundra", st_sfc(st_point(c(-165.43798, 64.52649))), wwfCentroid$geometry),
    crs = st_crs(wwfCentroid$geometry))
```

```{r}
wwfCentroid
```
## Reproject and finalize
```{r}
( wwfCentroidreproj = wwfCentroid %>% 
    st_transform(., crs = polarProj) %>% 
    dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]) %>%
  left_join(., wwfPtSum, by='wwfName') %>%
  mutate(prct = signif(percent,digits=3))
  
)


```

## Save WWF Centroids
```{r eval=FALSE, include=FALSE}
st_write(wwfCentroidreproj, "../outputs/spatialDataCreate/wwfEcoregionCentroid.shp", driver="ESRI Shapefile")
```


# Resolve Ecozones for Map

## Summaries sites by 

```{r}
(
  resPtCt <- resWwfPt %>% 
    st_drop_geometry() %>%
    group_by(resName) %>%
    summarise(count = n()) %>%
    mutate(total = sum(count)) %>%
  mutate(prop = round(count/total, 5)) %>%
  mutate(percent = round(prop*100, 2)) %>%
  mutate(prctLabel = ifelse(percent == 0.40, 0.4,
                ifelse(percent == 44.81, 44.8,
                ifelse(percent == 7.27, 7.3,
                ifelse(percent == 0.08, 0.1,
                ifelse(percent == 1.00, 1.0,
                ifelse(percent == 0.35, 0.4,
                ifelse(percent == 18.09, 18.1,
                ifelse(percent == 1.74, 1.7,
                ifelse(percent == 16.92, 16.9,
                ifelse(percent == 1.38, 1.4,
                ifelse(percent == 1.67, 1.7,
                ifelse(percent == 6.28, 6.3,
                ifelse(percent == 0.01, 0.01, -99))))))))))))))
  )
```
```{r}
( resPtSum <- resWwfPt %>%
                group_by(resName) %>%
                summarise(st_union(geometry)) %>%
                st_centroid() %>%
  left_join(., resPtCt, by= 'resName') %>%
  mutate(total = sum(count)) %>%
  mutate(prop = round(count/total, 5)) %>%
  mutate(percent = round(prop*100, 2)) %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry() %>%
    drop_na(resName) %>%
  mutate(prctLabel = ifelse(percent == 0.40, 0.4,
                ifelse(percent == 44.81, 44.8,
                ifelse(percent == 7.27, 7.3,
                ifelse(percent == 0.08, 0.1,
                ifelse(percent == 1.00, 1.0,
                ifelse(percent == 0.35, 0.4,
                ifelse(percent == 18.09, 18.1,
                ifelse(percent == 1.74, 1.7,
                ifelse(percent == 16.92, 16.9,
                ifelse(percent == 1.38, 1.4,
                ifelse(percent == 1.67, 1.7,
                ifelse(percent == 6.28, 6.3,
                ifelse(percent == 0.01, 0.01, -99))))))))))))))
  )

```

```{r}
write_csv(resPtSum, '../outputs/resCentroidSummary.csv')
```

## Centroid by each region
Fore the overview map
### 1. "Arctic coastal tundra"
```{r}
res01 =  resTunBor %>% 
    filter(resName=='Arctic coastal tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Arctic coastal tundra')
```

### 2. "Arctic foothills tundra"
```{r}
res02 =  resTunBor %>% 
    filter(resName=='Arctic foothills tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Arctic foothills tundra')
```

### 3. "Beringia lowland tundra"
```{r}
res03 =  resTunBor %>% 
    filter(resName=='Beringia lowland tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Beringia lowland tundra')
```


3              Beringia lowland tundra POINT (-160.8539 60.86536)
4               Beringia upland tundra POINT (-161.7837 62.63477)

### 4. "Beringia upland tundra" 
```{r}
( res04 =  resTunBor %>% 
    filter(resName=='Beringia upland tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Beringia upland tundra') )
```
Longitude: -165.43798
Latitude: 64.52649
### 5. Chukchi Peninsula tundra 
```{r}
res05 =  resTunBor %>% 
    filter(resName=="Chukchi Peninsula tundra") %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName="Chukchi Peninsula tundra")
```

### 6. "East Siberian taiga" 
```{r}
res06 =  resTunBor %>% 
    filter(resName=="East Siberian taiga") %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName="East Siberian taiga")
```


### 7. "Interior Alaska-Yukon lowland taiga"
```{r}
res07 =  resTunBor %>% 
    filter(resName=='Interior Alaska-Yukon lowland taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Interior Alaska-Yukon lowland taiga')
```

### 8. "Interior Yukon-Alaska alpine tundra"
```{r}
res08 =  resTunBor %>% 
    filter(resName=='Interior Yukon-Alaska alpine tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Interior Yukon-Alaska alpine tundra')
```

### 9. "Muskwa-Slave Lake forests"
```{r}
( res09 =  resTunBor %>% 
    filter(resName=='Muskwa-Slave Lake taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Muskwa-Slave Lake taiga') 
  )
```

### 10. "Northeast Siberian taiga"
```{r}
res10=  resTunBor %>% 
    filter(resName=='Northeast Siberian taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Northeast Siberian taiga')
```

### 11. "Northern Canadian Shield taiga"
```{r}
res11=  resTunBor %>% 
    filter(resName=='Northern Canadian Shield taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Northern Canadian Shield taiga')
```
### 12. "Northwest Territories taiga"
```{r}
( res12=  resTunBor %>% 
    filter(resName=='Northwest Territories taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName='Northwest Territories taiga')
)
```

### 13. "Yamal-Gydan tundra"
```{r}
res13 =  resTunBor %>% 
    filter(resName=="Yamal-Gydan tundra") %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(resName="Yamal-Gydan tundra")
```

## Recombine WWF ecoregions
```{r}
( 
  resCentroid = bind_rows(res01, res02, res03, res04, res05, res06, res07, res08, res09, res10, res11, res12, res13) %>% 
    rename(geometry = x)
)
```
 
## Edit geometry of 'Arctic coastal tundra' becaus of overalp
```{r}
resCentroid
```



-142.3674 70.75167
```{r eval=FALSE, include=FALSE}
st_geometry(resCentroid) <- st_sfc(
    ifelse(resCentroid$resName == 'Arctic coastal tundra', st_sfc(st_point(c(-142.3674, 70.75167))), resCentroid$geometry),
    crs = st_crs(resCentroid$geometry))
```

```{r}
resCentroid
```
## Reproject and finalize
```{r}
( resCentroidreproj = resCentroid %>% 
    st_transform(., crs = polarProj) %>% 
    dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]) %>%
  left_join(., resPtSum, by='resName') %>%
  mutate(prct = signif(percent,digits=3))
  
)


```

## Save WWF Centroids
```{r eval=FALSE, include=FALSE}
st_write(resCentroidreproj, "../outputs/spatialDataCreate/resEcozoneCentroid.shp", driver="ESRI Shapefile")
```


